---
title: "Using Shiny's Native OpenTelemetry with Bidux"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Shiny's Native OpenTelemetry with Bidux}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

OpenTelemetry is an industry-standard observability framework that provides distributed tracing, metrics, and logging capabilities. Starting with version 1.12.0, Shiny includes native OpenTelemetry support, allowing you to collect rich performance and interaction data from your applications without additional packages.

The `{bidux}` package now supports analyzing OpenTelemetry data alongside traditional `{shiny.telemetry}` data, providing a seamless workflow for UX friction detection regardless of your telemetry source.

### What is OpenTelemetry?

OpenTelemetry (OTEL) is a vendor-neutral, open-source observability framework that provides:

- **Distributed tracing**: Track requests and operations across services
- **Rich performance data**: Capture detailed timing information for renders, reactive updates, and user interactions
- **Standard format**: OTLP (OpenTelemetry Protocol) is widely supported by observability platforms
- **Flexible exports**: Send data to files, databases, or live collectors like Jaeger, Grafana, or Datadog

### Why Use OpenTelemetry with Shiny?

**Benefits of Shiny's native OpenTelemetry:**

1. **Built into Shiny** - No need to add telemetry packages to your application code
2. **Richer data** - Captures performance metrics (render times, reactive latency) alongside user interactions
3. **Modern observability** - Integrates with industry-standard monitoring tools
4. **Production-ready** - Designed for enterprise observability workflows

**When to use it:**

- You're running Shiny 1.12.0 or later
- You want detailed performance insights (not just interaction tracking)
- You need integration with existing observability infrastructure
- You're deploying production applications at scale

## Prerequisites

To use OpenTelemetry with Shiny and analyze the data with bidux, you need:

```{r prerequisites}
# 1. Shiny >= 1.12.0 (when OTEL support was added)
packageVersion("shiny") # Should be >= 1.12.0

# 2. OpenTelemetry packages
install.packages("otel")
install.packages("otelsdk")

# 3. bidux with OTEL support
install.packages("bidux") # or development version from GitHub
```

## Setting Up OpenTelemetry in Your Shiny App

### Basic Configuration

The simplest way to enable OpenTelemetry in your Shiny app is through options or environment variables:

```{r basic_setup}
library(shiny)

# Enable OTEL collection for all events
options(shiny.otel.collect = "all")

# Or use environment variable (set before starting R)
# Sys.setenv(SHINY_OTEL_COLLECT = "all")

# Your Shiny app code
ui <- fluidPage(
  titlePanel("Sales Dashboard"),
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Region:",
                  choices = c("North", "South", "East", "West")),
      dateRangeInput("date_range", "Date Range:")
    ),
    mainPanel(
      plotOutput("sales_plot")
    )
  )
)

server <- function(input, output, session) {
  output$sales_plot <- renderPlot({
    # Your plotting logic
  })
}

shinyApp(ui, server)
```

### Collection Levels

Shiny's OpenTelemetry supports different collection levels to control data volume:

| Level | What's Collected | Use Case |
|-------|-----------------|----------|
| `"none"` | No telemetry | Production (no monitoring) |
| `"session"` | Session start/end only | Minimal overhead tracking |
| `"reactive_update"` | Session + reactive updates | Balance of data and performance |
| `"reactivity"` | Above + reactive dependencies | Detailed reactive graph insights |
| `"all"` | Everything (max detail) | Development and analysis |

```{r collection_levels}
# For development and UX analysis - collect everything
options(shiny.otel.collect = "all")

# For production with moderate overhead
options(shiny.otel.collect = "reactive_update")

# Minimal production tracking
options(shiny.otel.collect = "session")
```

### Export Formats

#### Option 1: File Export (for Bidux Analysis)

To analyze OTEL data with bidux, export to OTLP JSON format:

```{r file_export}
\dontrun{
library(otel)
library(otelsdk)

# Configure OTLP JSON file exporter
Sys.setenv(
  OTEL_TRACES_EXPORTER = "otlp",
  OTEL_EXPORTER_OTLP_PROTOCOL = "http/json",
  OTEL_EXPORTER_OTLP_ENDPOINT = "file:///path/to/otel_spans.json"
)

# Enable collection
options(shiny.otel.collect = "all")

# Run your Shiny app
# Spans will be exported to otel_spans.json
}
```

#### Option 2: SQLite Database Export

For persistent storage compatible with bidux:

```{r sqlite_export}
\dontrun{
library(otel)
library(otelsdk)

# Configure SQLite exporter (custom implementation)
# Note: Requires additional setup - see otel package documentation
Sys.setenv(
  OTEL_TRACES_EXPORTER = "sqlite",
  OTEL_EXPORTER_SQLITE_PATH = "/path/to/otel_spans.sqlite"
)

options(shiny.otel.collect = "all")
}
```

#### Option 3: Live Collectors (Jaeger, Grafana, etc.)

For real-time monitoring in production:

```{r live_export}
\dontrun{
library(otel)
library(otelsdk)

# Configure OTLP HTTP exporter to send to collector
Sys.setenv(
  OTEL_TRACES_EXPORTER = "otlp",
  OTEL_EXPORTER_OTLP_PROTOCOL = "http/protobuf",
  OTEL_EXPORTER_OTLP_ENDPOINT = "https://collector.example.com:4318",
  OTEL_EXPORTER_OTLP_HEADERS = "Authorization=Bearer YOUR_TOKEN",
  OTEL_RESOURCE_ATTRIBUTES = "service.name=my-shiny-app,environment=production"
)

options(shiny.otel.collect = "all")
}
```

### Environment Variables Reference

Key environment variables for configuring OpenTelemetry:

```{r env_vars}
# Core configuration
SHINY_OTEL_COLLECT = "all"              # Collection level
OTEL_TRACES_EXPORTER = "otlp"           # Exporter type

# OTLP configuration
OTEL_EXPORTER_OTLP_ENDPOINT = "https://collector:4318"  # Collector URL
OTEL_EXPORTER_OTLP_PROTOCOL = "http/json"               # Protocol format
OTEL_EXPORTER_OTLP_HEADERS = "Authorization=Bearer token"  # Auth headers

# Resource attributes (metadata)
OTEL_RESOURCE_ATTRIBUTES = "service.name=my-app,environment=prod,version=1.0"

# Sampling (control data volume)
OTEL_TRACES_SAMPLER = "traceidratio"    # Sampler type
OTEL_TRACES_SAMPLER_ARG = "0.1"         # Sample 10% of traces
```

## How OTEL Spans Are Converted to Bidux Events

Before analyzing OTEL data, bidux automatically converts OpenTelemetry spans to its event schema. Understanding this conversion helps you:

1. Configure your Shiny app to emit useful spans
2. Understand what the analysis results mean
3. Debug issues when data doesn't look right

### Span Type Mappings

Bidux recognizes these OTEL span types and converts them to event types:

| Span Name/Pattern | Event Type | Description |
|-------------------|------------|-------------|
| `session_start` | `login` | User session begins |
| `session_end` | `logout` | User session ends |
| `output:<id>` | `output` | Shiny output rendering (e.g., `output:plot1`) |
| `reactive:<id>` | `input` | Reactive expression execution |
| `observe:<id>` | `input` | Observer execution |
| `navigation` | `navigation` | Tab/page navigation event |
| `reactive_update` | `reactive_update` | Reactive recalculation (timing event) |
| Span with error event | `error` | Error occurred during execution |

### Column Schema

After conversion, events have this schema:

| Column | Type | Description | Example |
|--------|------|-------------|---------|
| `timestamp` | POSIXct | Event time (from `startTimeUnixNano`) | `2025-01-15 14:23:01 UTC` |
| `session_id` | character | Session identifier (from `session.id` attribute) | `"session_abc123"` |
| `event_type` | character | Type of event | `"output"`, `"input"`, `"login"` |
| `input_id` | character | Input/reactive identifier | `"slider1"`, `"filtered_data"` |
| `output_id` | character | Output identifier | `"plot1"`, `"table1"` |
| `error_message` | character | Error description (if error occurred) | `"object not found"` |
| `navigation_id` | character | Navigation target | `"settings_tab"` |
| `duration_ms` | numeric | Span duration in milliseconds | `234.5` |
| `value` | character | Value (usually NA for OTEL) | `NA` |

### ID Extraction Logic

Bidux uses flexible extraction to handle various naming conventions:

**For Input IDs** (reactive and observer spans):
1. Check span name for patterns:
   - `reactive:input$<id>` → extracts `<id>`
   - `reactive:<id>` → extracts `<id>`
   - `observe:<id>` → extracts `<id>`
2. Check attributes for: `input_id`, `widget_id`, `element_id`

**For Output IDs** (output spans):
1. Check span name: `output:<id>` → extracts `<id>`
2. Check attributes for: `output_id`, `target_id`, `output`, `output.name`

**For Session IDs** (all spans):
- Check attributes for: `session.id`, `session_id`

**For Navigation IDs** (navigation spans):
- Check attributes for: `navigation_id`, `navigation.target`, `page`, `target`

**For Error Messages**:
- Look in span events for events named `error` or `exception`
- Extract from event attributes: `message`, `error.message`, `exception.message`

### Timestamp Conversion

OTLP uses Unix nanosecond timestamps. Bidux converts them:

```r
# OTLP timestamp example: "1704459200000000000" (nanoseconds since epoch)
# Converted to: 2024-01-05 12:00:00 UTC (POSIXct)

# Conversion formula:
timestamp_seconds <- as.numeric(startTimeUnixNano) / 1e9
timestamp_posix <- as.POSIXct(timestamp_seconds, origin = "1970-01-01", tz = "UTC")
```

### Duration Calculation

Span duration is calculated from start and end timestamps:

```r
# Example span:
# startTimeUnixNano: "1704459200000000000"
# endTimeUnixNano:   "1704459200234500000"

# Duration calculation:
duration_ms <- (endTimeUnixNano - startTimeUnixNano) / 1e6
# Result: 234.5 milliseconds
```

This gives you precise millisecond-level timing for:
- Output render times
- Reactive execution times
- Observer execution times

## Analyzing OTEL Data with Bidux

### The Workflow (Same as shiny.telemetry!)

The beauty of bidux's OTEL support is that **after conversion**, the analysis workflow is **identical** to `shiny.telemetry`:

```{r bidux_workflow}
library(bidux)
library(dplyr)

# Works just like shiny.telemetry!
issues <- bid_telemetry("otel_spans.json")

# Same friction detection
critical_issues <- issues |>
  filter(severity == "critical") |>
  arrange(desc(impact_rate))

# Same BID pipeline
interpret <- bid_interpret(
  central_question = "How to improve user experience based on OTEL data?"
)

notices <- bid_notices(
  issues = critical_issues,
  previous_stage = interpret,
  max_issues = 3
)

# Extract telemetry flags
flags <- bid_flags(issues)
flags$has_critical_issues
```

### Format Auto-Detection

Bidux automatically detects whether your data is from `shiny.telemetry` or OpenTelemetry:

```{r auto_detect}
# Automatically detects shiny.telemetry format
issues_st <- bid_telemetry("telemetry.sqlite")

# Automatically detects OTLP JSON format
issues_otel <- bid_telemetry("otel_spans.json")

# Automatically detects OTEL SQLite format
issues_otel_db <- bid_telemetry("otel_spans.sqlite")

# Same analysis, same results, regardless of source!
```

### Understanding OTEL Span Conversion

When bidux analyzes OTEL data, spans are automatically converted to the bidux event schema. Here's how the conversion works:

### Span Name Patterns

OTEL spans use naming conventions that bidux recognizes:

- **Session lifecycle**: `session_start`, `session_end`
- **Outputs**: `output:plot1`, `output:table1` (pattern: `output:<id>`)
- **Reactives**: `reactive:input$slider`, `reactive:filtered_data` (pattern: `reactive:<id>`)
- **Observers**: `observe:update_db` (pattern: `observe:<id>`)
- **Navigation**: `navigation` (target extracted from attributes)

### Attribute Extraction

Bidux extracts metadata from span attributes using multiple naming conventions:

**Session ID**: Looks for `session.id` or `session_id` attributes

**Input ID**: Looks for:
- `input_id` in attributes
- `input$<id>` pattern in span name
- `reactive:<id>` or `observe:<id>` patterns

**Output ID**: Looks for:
- `output_id`, `target_id`, `output`, or `output.name` in attributes
- `output:<id>` pattern in span name

**Navigation ID**: Looks for:
- `navigation_id`, `navigation.target`, `page`, or `target` attributes

**Error Messages**: Extracted from span events with name `error` or `exception`:
- Looks for `message`, `error.message`, or `exception.message` attributes

### Duration Calculation

OTEL spans include precise timing information:

```{r otel_duration}
# Duration calculated from span timestamps
# duration_ms = (endTimeUnixNano - startTimeUnixNano) / 1e6

# Analyze OTEL data
issues <- bid_telemetry("otel_spans.json")

# OTEL data provides performance context
issues |>
  filter(issue_type == "delayed_interaction") |>
  select(problem, evidence)
#> Problem: Users take a long time before making their first interaction
#> Evidence: Median time to first input is 47 seconds
```

### Complete Example: From Setup to Analysis

Here's a full workflow from configuring OTEL in your Shiny app to analyzing results with bidux:

```{r complete_example}
\dontrun{
# ============================================
# STEP 1: Configure OTEL in your Shiny app
# ============================================

library(shiny)
library(otel)
library(otelsdk)

# Enable OTEL with file export
Sys.setenv(
  OTEL_TRACES_EXPORTER = "otlp",
  OTEL_EXPORTER_OTLP_ENDPOINT = "file:///tmp/shiny_otel.json"
)
options(shiny.otel.collect = "all")

# Your Shiny app
ui <- fluidPage(
  titlePanel("Sales Dashboard"),
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Region:",
                  choices = c("North", "South", "East", "West")),
      selectInput("product", "Product:",
                  choices = c("A", "B", "C")),
      dateRangeInput("dates", "Date Range:")
    ),
    mainPanel(
      tabsetPanel(
        tabPanel("Overview", plotOutput("overview")),
        tabPanel("Details", tableOutput("details")),
        tabPanel("Settings", uiOutput("settings"))
      )
    )
  )
)

server <- function(input, output, session) {
  output$overview <- renderPlot({
    # Plotting logic
  })

  output$details <- renderTable({
    # Table logic
  })

  output$settings <- renderUI({
    # Settings UI
  })
}

# Run app and collect data
shinyApp(ui, server)

# ============================================
# STEP 2: Analyze OTEL data with bidux
# ============================================

library(bidux)
library(dplyr)

# Analyze collected OTEL spans
issues <- bid_telemetry(
  "/tmp/shiny_otel.json",
  thresholds = bid_telemetry_presets("moderate")
)

# Review identified issues
print(issues)
#> # BID Telemetry Issues Summary
#> Found 5 issues from 342 sessions
#>
#> Critical: 1 issue
#> High: 2 issues
#> Medium: 2 issues

# Filter to critical issues
critical <- issues |>
  filter(severity == "critical")

print(critical)
#> Issue: unused_input_product
#> Problem: Users are not interacting with the 'product' input control
#> Evidence: Only 12 out of 342 sessions (3.5%) interacted with 'product'
#> Impact: 96.5% of sessions affected

# ============================================
# STEP 3: Apply BID framework
# ============================================

# Start BID workflow with OTEL insights
interpret_result <- bid_interpret(
  central_question = "Why aren't users engaging with the product filter?",
  data_story = new_data_story(
    hook = "96.5% of users never use the product filter",
    context = "OTEL data from 342 sessions over 2 weeks",
    tension = "Filter may be unnecessary or poorly positioned",
    resolution = "Simplify interface or improve filter discoverability"
  )
)

# Convert OTEL issue to Notice
notice_result <- bid_notices(
  issues = critical,
  previous_stage = interpret_result
)[[1]]

# Continue through BID stages
anticipate_result <- bid_anticipate(
  previous_stage = notice_result,
  bias_mitigations = list(
    choice_overload = "Reduce number of visible filters",
    default_effect = "Set smart defaults based on common patterns"
  )
)

# Use OTEL flags to inform structure
flags <- bid_flags(issues)
structure_result <- bid_structure(
  previous_stage = anticipate_result,
  telemetry_flags = flags
)

# Validate
validate_result <- bid_validate(
  previous_stage = structure_result,
  summary_panel = "Simplified dashboard with progressive disclosure",
  next_steps = c(
    "Remove or hide unused product filter",
    "Re-run OTEL analysis to verify improvement",
    "Monitor user engagement metrics"
  )
)

# Generate report
bid_report(validate_result, format = "html")
}
```

## Comparison: shiny.telemetry vs Shiny Native OTEL

Understanding the differences helps you choose the right approach:

| Feature | shiny.telemetry | Shiny OTEL |
|---------|----------------|------------|
| **Setup** | Separate package | Built into Shiny 1.12+ |
| **In-app code** | `use_telemetry()` + tracking calls | Just set `options(shiny.otel.collect)` |
| **Data captured** | User interactions | Interactions + performance spans |
| **Format** | Events (direct) | Spans (converted to events by bidux) |
| **Performance data** | Limited (manual timing) | Rich (automatic render times, reactive latency) |
| **File size** | Smaller (event-only) | Larger (includes span metadata) |
| **Shiny version** | Works with older Shiny | Requires Shiny >= 1.12.0 |
| **Bidux support** | Yes (native) | Yes (automatic conversion) |
| **Best for** | Simple tracking, older Shiny | Performance insights, modern Shiny |

### When to Use Each

**Use shiny.telemetry when:**

- Using Shiny versions < 1.12.0
- You need lightweight event tracking only
- You have an established `shiny.telemetry` workflow
- File size and storage are constraints
- You want fine-grained control over what's tracked

**Use Shiny OpenTelemetry when:**

- Using Shiny >= 1.12.0
- You want comprehensive performance insights
- You're integrating with existing OTEL infrastructure
- You need distributed tracing across services
- You want automatic tracking without instrumentation code

**Use both during transition:**

You can run both systems simultaneously during migration:

```{r dual_tracking}
\dontrun{
library(shiny)
library(shiny.telemetry)
library(otel)

# Enable both systems
telemetry <- Telemetry$new()
options(shiny.otel.collect = "all")

ui <- fluidPage(
  use_telemetry(),  # shiny.telemetry
  # Your UI
)

server <- function(input, output, session) {
  telemetry$start_session()  # shiny.telemetry
  # Your server logic
}

# Analyze both sources
issues_st <- bid_telemetry("telemetry.sqlite")
issues_otel <- bid_telemetry("otel_spans.json")

# Compare results
nrow(issues_st)
nrow(issues_otel)
}
```

## Migration Considerations

### Should You Switch from shiny.telemetry to OTEL?

**Reasons to migrate:**

1. **Automatic instrumentation** - No need to add tracking code
2. **Richer data** - Performance metrics included automatically
3. **Standard format** - OTLP is widely supported
4. **Future-proof** - OTEL is the industry standard

**Reasons to stay with shiny.telemetry:**

1. **Shiny version** - You're on Shiny < 1.12.0
2. **Simplicity** - You only need basic event tracking
3. **Storage** - OTEL data is more verbose
4. **Established workflow** - You have working pipelines

### Migration Strategy

If you decide to migrate, here's a phased approach:

**Phase 1: Parallel tracking (2-4 weeks)**

```{r migration_phase1}
\dontrun{
# Run both systems to compare
options(shiny.otel.collect = "all")
# Keep existing shiny.telemetry code

# Compare results weekly
issues_st <- bid_telemetry("telemetry.sqlite")
issues_otel <- bid_telemetry("otel_spans.json")

# Verify OTEL captures same issues
}
```

**Phase 2: OTEL primary (2-4 weeks)**

```{r migration_phase2}
\dontrun{
# Switch to OTEL as primary
options(shiny.otel.collect = "all")
# Keep shiny.telemetry as backup

# Use OTEL data for analysis
issues <- bid_telemetry("otel_spans.json")
}
```

**Phase 3: OTEL only**

```{r migration_phase3}
\dontrun{
# Remove shiny.telemetry code
# library(shiny.telemetry) - remove
# use_telemetry() - remove
# telemetry$start_session() - remove

# OTEL only
options(shiny.otel.collect = "all")
}
```

## Troubleshooting

### Common Issues and Solutions

**Problem: "otel package not found"**

```{r troubleshoot_1}
# Solution: Install OpenTelemetry packages
install.packages("otel")
install.packages("otelsdk")
```

**Problem: "No spans detected"**

```{r troubleshoot_2}
# Check if OTEL is enabled
getOption("shiny.otel.collect")
#> Should return "all" or another collection level

# Verify otel is tracing
library(otel)
otel::is_tracing_enabled()
#> Should return TRUE

# Enable OTEL if disabled
options(shiny.otel.collect = "all")
```

**Problem: "Format not recognized" when analyzing OTLP JSON**

```{r troubleshoot_3}
# Verify file structure
jsonlite::fromJSON("otel_spans.json", simplifyVector = FALSE) |>
  str(max.level = 2)

# Should contain spans with startTimeUnixNano, endTimeUnixNano, etc.
# If not, check OTLP exporter configuration
```

**Problem: "Empty spans file"**

```{r troubleshoot_4}
# Check exporter endpoint
Sys.getenv("OTEL_EXPORTER_OTLP_ENDPOINT")

# Verify file path is writable
file.access("otel_spans.json", mode = 2)
#> Should return 0 (success)

# Check Shiny app had user interactions
# Spans only created when actions occur
```

**Problem: "Too much data / large files"**

```{r troubleshoot_5}
# Use sampling to reduce volume
Sys.setenv(
  OTEL_TRACES_SAMPLER = "traceidratio",
  OTEL_TRACES_SAMPLER_ARG = "0.1"  # Sample 10% of traces
)

# Or reduce collection level
options(shiny.otel.collect = "reactive_update")  # Less than "all"
```

**Problem: "Bidux not detecting OTEL format"**

```{r troubleshoot_6}
# Explicitly specify format
issues <- bid_telemetry("otel_spans.json", format = "otlp_json")

# Or for OTEL SQLite
issues <- bid_telemetry("otel_spans.sqlite", format = "otel_sqlite")
```

## Advanced Topics

### Custom Span Attributes

You can add custom attributes to OTEL spans for richer analysis:

```{r custom_attributes}
\dontrun{
library(otel)

# Add custom attributes to current span
otel::add_span_attribute("user_role", "analyst")
otel::add_span_attribute("dashboard_version", "2.1.0")

# These attributes are preserved in OTLP exports
# and available for custom analysis
}
```

### Filtering Analysis by Attributes

```{r filter_attributes}
\dontrun{
# Analyze OTEL data
issues <- bid_telemetry("otel_spans.json")

# Access raw span data for custom filtering
# (Advanced: requires understanding OTLP structure)
raw_spans <- jsonlite::fromJSON("otel_spans.json")

# Filter spans by custom attributes before analysis
# Then re-analyze with bidux
}
```

## Best Practices

1. **Start with "all" during development** - Collect everything for UX analysis
   ```{r bp1}
   options(shiny.otel.collect = "all")
   ```

2. **Use sampling in production** - Reduce overhead with sampling
   ```{r bp2}
   Sys.setenv(OTEL_TRACES_SAMPLER_ARG = "0.1")  # 10% sampling
   ```

3. **Rotate log files** - Prevent unbounded file growth
   ```{r bp3}
   \dontrun{
   # Implement log rotation in your deployment
   # Example: daily rotation with retention
   file_pattern <- paste0("otel_spans_", Sys.Date(), ".json")
   }
   ```

4. **Monitor file sizes** - OTEL data can grow quickly
   ```{r bp4}
   \dontrun{
   file.size("otel_spans.json") / 1024 / 1024  # Size in MB
   }
   ```

5. **Regular analysis** - Run bidux analysis weekly or monthly
   ```{r bp5}
   # Schedule regular UX reviews
   issues <- bid_telemetry("otel_spans.json")
   if (any(issues$severity == "critical")) {
     # Alert team
   }
   ```

6. **Combine with user feedback** - OTEL shows *what*, interviews show *why*
   ```{r bp6}
   # Use OTEL to identify friction points
   # Then interview users to understand root causes
   ```

## References

- [Shiny Changelog - Version 1.12.0](https://rstudio.github.io/shiny/news/index.html)
- [otel Package Documentation](https://otel.r-lib.org/)
- [otelsdk Package Documentation](https://otelsdk.r-lib.org/)
- [OpenTelemetry Specification](https://opentelemetry.io/docs/specs/otel/)
- [OTLP Protocol Reference](https://opentelemetry.io/docs/specs/otlp/)
- [bidux Package](https://github.com/jrwinget/bidux)

## Next Steps

Now that you understand OpenTelemetry integration with bidux:

1. **Set up OTEL** in your Shiny app following the configuration examples
2. **Collect data** from real users (at least 50-100 sessions)
3. **Analyze with bidux** using `bid_telemetry()`
4. **Apply BID framework** to address identified friction points
5. **Measure improvement** by comparing before/after metrics

For more details on the BID framework and telemetry analysis:

- `vignette("telemetry-integration")` - Comprehensive telemetry guide
- `vignette("getting-started")` - BID framework introduction
- `vignette("practical-examples")` - Real-world use cases

Happy analyzing!
