---
title: "Integrating shiny.telemetry with BIDUX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Integrating shiny.telemetry with BIDUX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `bid_ingest_telemetry()` function enables you to analyze real user behavior data from `{shiny.telemetry}` and automatically identify UX friction points. This creates a powerful feedback loop where actual usage patterns drive design improvements through the BID framework.

## Prerequisites

First, ensure you have `{shiny.telemetry}` set up in your Shiny application:

```{r setup, eval=FALSE}
library(shiny)
library(shiny.telemetry)

# Initialize telemetry
telemetry <- Telemetry$new()

ui <- fluidPage(
  use_telemetry(), # Add telemetry JavaScript

  titlePanel("Sales Dashboard"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        "region",
        "Region:",
        choices = c("North", "South", "East", "West")
      ),
      dateRangeInput("date_range", "Date Range:"),
      selectInput(
        "product_category",
        "Product Category:",
        choices = c("All", "Electronics", "Clothing", "Food")
      ),
      actionButton("refresh", "Refresh Data")
    ),
    mainPanel(
      tabsetPanel(
        tabPanel("Overview", plotOutput("overview_plot")),
        tabPanel("Details", dataTableOutput("details_table")),
        tabPanel("Settings", uiOutput("settings_ui"))
      )
    )
  )
)

server <- function(input, output, session) {
  # Start telemetry tracking
  telemetry$start_session()

  # Your app logic here...
}

shinyApp(ui, server)
```

## Analyzing Telemetry Data

After collecting telemetry data from your users, use `bid_ingest_telemetry()` to identify UX issues:

```{r basic_usage, eval=FALSE}
library(bidux)

# Analyze telemetry from SQLite database (default)
issues <- bid_ingest_telemetry("telemetry.sqlite")

# Or from JSON log file
issues <- bid_ingest_telemetry("telemetry.log", format = "json")

# Review identified issues
length(issues)
names(issues)
```

## Understanding the Analysis

The function analyzes five key friction indicators:

### 1. Unused or Under-used Inputs

Identifies UI controls that users rarely or never interact with:

```{r unused_inputs, eval=FALSE}
# Example: Region filter is never used
issues$unused_input_region
#> BID Framework - Notice Stage
#> Problem: Users are not interacting with the 'region' input control
#> Theory: Hick's Law (auto-suggested)
#> Evidence: Telemetry shows 0 out of 847 sessions where 'region' was changed
```

This suggests the region filter might be:
- Hidden or hard to find
- Not relevant to users' tasks
- Confusing or intimidating

### 2. Delayed First Interactions

Detects when users take too long to start using the dashboard:

```{r delayed_interaction, eval=FALSE}
issues$delayed_interaction
#> BID Framework - Notice Stage
#> Problem: Users take a long time before making their first interaction with the dashboard
#> Theory: Information Scent (auto-suggested)
#> Evidence: Median time to first input is 45 seconds, and 10% of sessions had no interactions at all
```

This indicates users might be:
- Overwhelmed by the initial view
- Unsure where to start
- Looking for information that's not readily apparent

### 3. Frequent Errors

Identifies systematic errors that disrupt user experience:

```{r error_patterns, eval=FALSE}
issues$error_1
#> BID Framework - Notice Stage
#> Problem: Users encounter errors when using the dashboard
#> Theory: Norman's Gulf of Evaluation (auto-suggested)
#> Evidence: Error 'Data query failed' occurred 127 times in 15.0% of sessions (in output 'overview_plot'), often after changing 'date_range'
```

This reveals:
- Reliability issues with specific features
- Input validation problems
- Performance bottlenecks

### 4. Navigation Drop-offs

Finds pages or tabs that users rarely visit:

```{r navigation_dropoff, eval=FALSE}
issues$navigation_settings_tab
#> BID Framework - Notice Stage
#> Problem: The 'settings_tab' page/tab is rarely visited by users
#> Theory: Information Architecture (auto-suggested)
#> Evidence: Only 42 sessions (5.0%) visited 'settings_tab', and 90% of those sessions ended there
```

Low visit rates suggest:
- Poor information scent
- Hidden or unclear navigation
- Irrelevant content

### 5. Confusion Patterns

Detects rapid repeated changes indicating user confusion:

```{r confusion_pattern, eval=FALSE}
issues$confusion_date_range
#> BID Framework - Notice Stage
#> Problem: Users show signs of confusion when interacting with 'date_range'
#> Theory: Feedback Loops (auto-suggested)
#> Evidence: 8 sessions showed rapid repeated changes (avg 6 changes in 7.5 seconds), suggesting users are unsure about the input's behavior
```

This suggests:
- Unclear feedback when values change
- Unexpected behavior
- Poor affordances

## Customizing Analysis Thresholds

You can adjust the sensitivity of the analysis:

```{r custom_thresholds, eval=FALSE}
issues <- bid_ingest_telemetry(
  "telemetry.sqlite",
  thresholds = list(
    unused_input_threshold = 0.1, # Flag if <10% of sessions use input
    delay_threshold_seconds = 60, # Flag if >60s before first interaction
    error_rate_threshold = 0.05, # Flag if >5% of sessions have errors
    navigation_threshold = 0.3, # Flag if <30% visit a page
    rapid_change_window = 5, # Look for 5 changes within...
    rapid_change_count = 3 # ...3 seconds
  )
)
```

## Integrating with BID Workflow

Use the identified issues to drive your BID process:

```{r bid_workflow, eval=FALSE}
# Take the most critical issue
critical_issue <- issues$error_1

# Continue with interpretation
interpret_result <- bid_interpret(
  previous_stage = critical_issue,
  central_question = "How can we prevent data query errors?",
  data_story = list(
    hook = "15% of users encounter errors",
    context = "Errors occur after date range changes",
    tension = "Users lose trust when queries fail",
    resolution = "Implement robust error handling and loading states"
  )
)

# Structure improvements
structure_result <- bid_structure(
  previous_stage = interpret_result,
  concepts = c("Progressive Disclosure", "Feedback Loops")
)

# Continue through anticipate and validate stages...
```

## Example: Complete Analysis

Here's a full example analyzing a dashboard with multiple issues:

```{r complete_example, eval=FALSE}
# 1. Ingest telemetry
issues <- bid_ingest_telemetry("telemetry.sqlite")

# 2. Prioritize issues by impact
critical_issues <- list(
  unused_inputs = names(issues)[grepl("unused_input", names(issues))],
  errors = names(issues)[grepl("error", names(issues))],
  delays = "delayed_interaction" %in% names(issues)
)

# 3. Create a comprehensive improvement plan
if (length(critical_issues$unused_inputs) > 0) {
  # Address unused inputs
  unused_issue <- issues[[critical_issues$unused_inputs[1]]]

  improvement_plan <- bid_interpret(
    previous_stage = unused_issue,
    central_question = "Which filters actually help users find insights?"
  ) |>
    bid_structure(
      concepts = c("Progressive Disclosure", "Smart Defaults")
    ) |>
    bid_anticipate(
      bias_mitigations = list(
        choice_overload = "Hide advanced filters until needed",
        default_effect = "Pre-select most common filter values"
      )
    ) |>
    bid_validate(
      summary_panel = "Show only relevant filters based on user's task",
      next_steps = c(
        "Remove unused filters",
        "Implement progressive disclosure",
        "Add contextual help",
        "Re-test with telemetry after changes"
      )
    )
}

# 4. Generate report
improvement_report <- bid_report(improvement_plan, format = "html")
```

## Best Practices

1. **Collect Sufficient Data**: Ensure you have telemetry from at least 50-100 sessions before analysis for reliable patterns.

2. **Regular Analysis**: Run telemetry analysis periodically (e.g., monthly) to catch emerging issues.

3. **Combine with Qualitative Data**: Use telemetry insights alongside user interviews and usability testing.

4. **Track Improvements**: After implementing changes, collect new telemetry to verify improvements:

```{r verify_improvements, eval=FALSE}
# Before changes
issues_before <- bid_ingest_telemetry("telemetry_before.sqlite")

# After implementing improvements
issues_after <- bid_ingest_telemetry("telemetry_after.sqlite")

# Compare issue counts
cat("Issues before:", length(issues_before), "\n")
cat("Issues after:", length(issues_after), "\n")
```

5. **Document Patterns**: Build a knowledge base of common patterns in your domain:

```{r document_patterns, eval=FALSE}
# Save recurring patterns for future reference
telemetry_patterns <- list(
  date_filter_confusion = "Users often struggle with date range inputs - consider using presets",
  tab_discovery = "Secondary tabs have low discovery - consider better visual hierarchy",
  error_recovery = "Users abandon after errors - implement graceful error handling"
)
```

## Conclusion

The `bid_ingest_telemetry()` function bridges the gap between user behavior data and design decisions. By automatically identifying friction points from real usage patterns, it provides concrete, evidence-based starting points for the BID framework, ultimately leading to more user-friendly Shiny applications.
